\section*{Grading $SQL$ requirements}
\textbf{Must-have requirements:}
\begin{enumerate}[label=R-\arabic*]
  \item The library can connect to a $MySQL$ database when passed
    connection details (server location - port, IP, user - and database name)
  \item The library can handle any error that might occur when communicating to
    the database (including failed execution of SQL queries, connection issues, etc.)
  \item The library can perform $MySQL$ queries on a database.
  \item The library can clear all tables and data from a database.
  \item The library can compare the results of two $SQL$ queries and check if they
    match.
  \item The library can canonicalize two $SQL$ queries in order
    \textit{to remove remove any irrelevant syntactic variation} as much as
    possible.
  \cite{literature:xdata}
  \item The library can return differences in the column list, table list (including \texttt{JOIN}
    conditions), \texttt{WHERE} conditions, \texttt{HAVING} conditions and \texttt{GROUP BY} conditions.
  \item The library can return a partial grade based on the mistakes made
  by the user
\end{enumerate}

\textbf{Additional requirements:}
\begin{enumerate}[resume,label=R-OPT-\arabic*]
  \item The library can generate a data set when given a database schema
  and an instructor's query.
\end{enumerate}

\section*{Web application requirements}
\textbf{Must-have requirements:}
\begin{enumerate}[resume*,label=R-\arabic*]
  \item The web application allows users to log-in and keep their data
    across sessions.
  \item The web application allows users to have different roles within the
    application (e.g. Student/ Teacher/ TA)
  \item  A teacher can create assignments by giving a title, a description,
    and a $SQL$ query that creates the tables and the relations between them.
    Optionally, the teacher can also provide a query that inserts some
    initial data.
  \item A student can provide solutions to assignments by submitting a
  $SQL$ query. The web application returns the results to the students
  after it assess the query.
\end{enumerate}

\textbf{Additional requirements:}
\begin{enumerate}[resume,label=R-OPT-\arabic*]
  \item The web application allows the teacher to set a release date and a group
  of students to whom the assignment shall be given to.
  \item $SQL$ typing on the web application should be made in a browser based
  code editor that provides syntax highlighting and suggestions.
  \item Teachers should be able to have a graphical view of the results.
\end{enumerate}
